---
execute:
  eval: false
---

# Extra – Mean stats for actigraphy data

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

## Overview

This notebook aims to help compute mean statistics for actigraphy data.

## Setup

```{r}
#| eval: false

library(astroFns, quietly = TRUE)
library(checkmate, quietly = TRUE)
library(cli, quietly = TRUE)
library(circular, quietly = TRUE)
library(here, quietly = TRUE)
library(hms, quietly = TRUE)
library(lubritime, quietly = TRUE) # github.com/danielvartan/lubritime
library(rutils, quietly = TRUE) # github.com/danielvartan/rutils
```

```{r}
#| includes: false

library(magrittr)
```

## Importing data

```{r}
#| output: false

googlesheets4::gs4_auth()
```

```{r}
#| output: false

ss <- "1iqjhdSiWYr9RYSvIvJQfFw35DR31YA_oj2SW8iR5WPw"

raw_data <- ss |> googlesheets4::read_sheet(sheet = "Dataset")
```

## Tidying data

```{r}
parse_hms_as_duration <- function(x) {
  x |>
    hms::parse_hms() |>
    lubridate::as.duration()
}
```

```{r}
data <-
  raw_data |>
  dplyr::mutate(
    dplyr::across(
      .cols = dplyr::matches("^his|^hfs"),
      .fns = hms::parse_hms
    ),
    dplyr::across(
      .cols = dplyr::matches("^fps|^tts_m|^tts_s|^waso"),
      .fns = parse_hms_as_duration
    ),
    dplyr::across(
      .cols = dplyr::matches("^tts_fps|^awakenings"),
      .fns = as.numeric
    )
  )
```

```{r}
data |> dplyr::glimpse()
```

## Analysis

```{r}
hms_to_decimal_hours <- function(x) {
  rutils:::assert_hms(x)

  x |>
    as.numeric() %>%
    `/`(60 * 60)
}
```

```{r}
decimal_hours_to_hms <- function(x) {
  checkmate::assert_numeric(x)

  x %>%
    `*`(60 * 60) |>
    hms::as_hms()
}
```

```{r}
hms_to_circular <- function(
    x,
    ...,
    type = "angles",
    units = "hours",
    template = "clock24"
) {
  rutils:::assert_hms(x)

  x |>
    hms_to_decimal_hours() |>
    circular::circular(
      type = type,
      units = units,
      template = template,
      ...
    )
}
```

```{r}
circular_time_stat <- function(x, fun = circular::mean.circular) {
  rutils:::assert_hms(x)
  checkmate::assert_function(fun)

  if (identical(fun, circular::sd.circular) ||
      identical(fun, stats::sd)) {
    x |>
      hms_to_circular() |>
      fun() |>
      astroFns::rad2hms() |>
      hms::as_hms() |>
      hms::round_hms(secs = 1)
  } else {
    x |>
      hms_to_circular() |>
      fun() |>
      as.numeric() |>
      lubritime::cycle_time(cycle = 24) |>
      decimal_hours_to_hms() |>
      hms::round_hms(secs = 1)
  }
}
```

```{r}
alt_circular_time_stat <- function(x, fun = mean) {
  rutils:::assert_hms(x)
  checkmate::assert_function(fun)

  x |>
    lubritime:::link_to_timeline(threshold = hms::parse_hms("12:00:00")) |>
    fun() |>
    hms::as_hms() |>
    hms::round_hms(secs = 1)
}
```

```{r}
linear_time_stat <- function(x, fun = mean) {
  rutils:::assert_duration(x)
  checkmate::assert_function(fun)

  x |>
    fun() |>
    hms::as_hms() |>
    hms::round_hms(secs = 1)
}
```

```{r}
rounded_stat <- function(x, fun = mean) {
  checkmate::assert_numeric(x)
  checkmate::assert_function(fun)

  x |>
    fun() |>
    round(3)
}
```

### What is the aggregate mean/standasrd deviation of the actigraphic summary statistics from the selected pregnant women?

```{r}
aggregate_mean_data <-
  data |>
  dplyr::select(-id) |>
  dplyr::group_by() |>
  dplyr::summarise(
    dplyr::across(
      .cols = dplyr::matches("^his|^hfs"),
      .fns = ~ circular_time_stat(.x, fun = circular::mean.circular)
    ),
    dplyr::across(
      .cols = dplyr::matches("^fps|^tts_m|^tts_s|^waso"),
      .fns = ~ linear_time_stat(.x, fun = mean)
    ),
    dplyr::across(
      .cols = dplyr::matches("^tts_fps|^awakenings"),
      .fns = ~ rounded_stat(.x, fun = mean)
    )
  )|>
  tidyr::pivot_longer(
    cols = dplyr::everything(),
    names_to = c(".value", "type"),
    names_sep = "_(?!.*_)",
  ) |>
  dplyr::relocate(tts_fps, .after = tts)
```

```{r}
aggregate_mean_data
```

```{r}
aggregate_sd_data <-
  data |>
  dplyr::select(-id) |>
  dplyr::group_by() |>
  dplyr::summarise(
    dplyr::across(
      .cols = dplyr::matches("^his|^hfs"),
      .fns = ~ circular_time_stat(.x, fun = circular::sd.circular)
    ),
    dplyr::across(
      .cols = dplyr::matches("^fps|^tts_m|^tts_s|^waso"),
      .fns = ~ linear_time_stat(.x, fun = stats::sd)
    ),
    dplyr::across(
      .cols = dplyr::matches("^tts_fps|^awakenings"),
      .fns = ~ rounded_stat(.x, fun = stats::sd)
    )
  )|>
  tidyr::pivot_longer(
    cols = dplyr::everything(),
    names_to = c(".value", "type"),
    names_sep = "_(?!.*_)",
  ) |>
  dplyr::relocate(tts_fps, .after = tts)
```

```{r}
aggregate_sd_data
```

## Comparing methods

```{r}
circular_vector <- data |> magrittr::extract2("his_max")
```

### A primitive way for doing circular statistics

```{r}
x <- circular_vector |> hms_to_decimal_hours()
x <- ifelse(x < 1, x + 24, x)

x |>
  mean() |>
  decimal_hours_to_hms() |>
  hms::round_hms(secs = 1)
```

```{r}
x |>
  stats::sd() |>
  decimal_hours_to_hms() |>
  hms::round_hms(secs = 1)
```

### Using `lubritime:::link_to_timeline()` with mid-day as threshold

```{r}
circular_vector |> alt_circular_time_stat(fun = mean)
```

```{r}
circular_vector |> alt_circular_time_stat(fun = stats::sd)
```

### How much we lost by transformig the `hms` vector to decimal hours?

```{r}
dplyr::tibble(
  original = circular_vector,
  processed = circular_vector |>
    hms_to_decimal_hours() |>
    decimal_hours_to_hms()
)
```

### Circular statistics using the `circular` package

```{r}
circular_vector |> hms_to_circular()
```

```{r}
circular_vector |>
  hms_to_circular() |>
  plot()
```

```{r}
circular_vector |>
  hms_to_circular() |>
  circular::mean.circular()
```

```{r}
circular_vector |> circular_time_stat(fun = mean)
```

```{r}
circular_vector |>
  hms_to_circular() |>
  circular::sd.circular() |>
  astroFns::rad2hms() |>
  hms::as_hms() |>
  hms::round_hms(secs = 1)
```
