<!-- %:::% .common h1 begin %:::% -->
# The Expression of the Sleep-Wake Cycle Throughout Pregnancy
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

```{r}
#| echo: false
#| output: asis

quartor:::quarto_status(
  type = "drafting",
  of_what = "of this thesis",
  latex_parskip = "\\microskip"
  )
```

::: {.callout-note}
## Target journal

1. [Chronobiology International](https://www.tandfonline.com/action/authorSubmission?show=instructions&journalCode=icbi20) ([IF 2022: 2.8/JCR](https://jcr.clarivate.com/jcr/) | [A1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
2. [Journal of Biological Rhythms](https://journals.sagepub.com/author-instructions/JBR) ([IF 2022: 3.5/JCR](https://jcr.clarivate.com/jcr/) | [A2/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
:::

::: {.callout-note}
The following study was performed by Alícia Rafaelly Vilefort Sales (**ARVS**), Daniel Vartanian (**DV**), Maria Augusta Medeiros de Andrade (**MAMA**), Mario Pedrazzoli (**MP**) and Christiane Borges do Nascimento Chofakian (**CBNC**).

\microskip

**ARVS**, **DV** and **MAMA** contributed to the study's  design. **ARVS** implemented the study, performed the statistical analysis, and authored the manuscript. **CBNC** and **MP** served as scientific advisors. All authors participated in discussions about the results and contributed to the final manuscript revision.

\microskip

_Future reference_: Sales, A. R. V., Vartanian, D., Andrade, M. A. M., Pedrazzoli, M., & Chofakian, C. B. N. (2024). The expression of the sleep-wake cycle throughout pregnancy: a systematic and quantitative literature review. _Chronobiology International_.
:::
