---
execute:
  eval: false
---

# Preliminary Analysis 2

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

## Set the Environment

```{r}
library(actverse) # github.com/danielvartan/actverse
library(askpass)
library(broom)
library(checkmate)
library(cli)
library(dplyr)
library(effectsize)
library(fs)
library(ggplot2)
library(googlesheets4)
library(here)
library(hms)
library(infer)
library(lockr) # github.com/danielvartan/lokcr
library(lubridate)
library(lubritime) # github.com/danielvartan/lubritime
library(magrittr)
library(orbis) # github.com/danielvartan/orbis
library(osfr)
library(plotr) # github.com/danielvartan/plotr
library(pwr)
library(readr)
library(rutils) # github.com/danielvartan/rutils
library(scaler) # github.com/danielvartan/scaler
library(scales)
library(sodium)
library(stats)
library(stringr)
library(summarytools)
library(tidyr)
library(tsibble)
```

```{r}
source(here::here("R", "anonymize_id.R"))
source(here::here("R", "ga.R"))
```

## Setting Parameters

```{r}
#| eval: false
#| output: false

prettycheck:::assert_interactive()

googlesheets4::gs4_auth()
```

```{r}
#| eval: false

prettycheck:::assert_interactive()

osf_auth()
```

```{r}
private_key_path <- here::here("_ssh", "id_rsa")
```

```{r}
#| eval: false

password <- askpass::askpass()
```

```{r}
#| eval: false

salt <- askpass::askpass()
```

```{r}
public_key_path <- here::here("_ssh", "id_rsa.pub")
```

## Get Data from Google Sheets

```{r}
#| output: false

control_form <- googlesheets4::read_sheet(
  ss = "13wtDr4fRD1wJSM-qdLOdGSchF8oXU1bG0Jtccu24GhY",
  sheet = "Dataset",
  col_types = "c"
)
```

## Process the Google Sheets Data

```{r}
control_data <-
  control_form |>
  `names<-`(paste0("X", seq_along(control_form))) |>
  dplyr::select(X5, X6, X12, X39, X40) |>
  dplyr::rename(
    cpf = X5,
    id = X6,
    birth_center_delivery = X12,
    ultrasound_date = X40,
    ultrasound_ga = X39
  ) |>
  dplyr::mutate(
    birth_center_delivery = dplyr::case_when(
      tolower(trimws(birth_center_delivery)) == "sim" ~ TRUE,
      tolower(trimws(birth_center_delivery)) == "não" ~ FALSE,
      TRUE ~ NA
    ),
    ultrasound_date = ymd(ultrasound_date),
    ga_start = if_else(
      is.na(ultrasound_ga),
      as.Date(NA),
      ga_start(
        ultrasound = ultrasound_date,
        ga =
          str_extract(ultrasound_ga, "^\\d{1,2}") |>
          as.integer() |>
          weeks() |>
          add(
            str_extract(ultrasound_ga, "\\d(?= di)") |>
            as.integer() |>
            lubridate::days()
          )
      )
    )
  ) |>
  select(cpf, id, birth_center_delivery, ga_start)

control_data
```

## Getting Data from OSF

```{r}
data_dir <- here::here("data")
```

```{r}
if (!dir.exists(data_dir)) dir.create(data_dir)
```

```{r}
osf_processed_data_id <- "a2dsw"
```

```{r}
#| eval: false

osf_processed_data_id |>
  osf_retrieve_node() |>
  osf_ls_files(n_max = Inf) |>
  filter(name == "actigraphy") |>
  pull(id) |>
  osf_retrieve_file() |>
  osf_ls_files(n_max = Inf) |>
  filter(stringr::str_detect(name, "actigraphy-processed-data")) |>
  osf_download(
    path = data_dir,
    conflicts = "overwrite",
    progress = TRUE
  )
```

## Decrypt Data from OSF

```{r}
data_dir |>
  unlock_dir(
    private_key = private_key_path,
    password = password
  )
```

## Sample Characteristics

## Import and Transform the Data

```{r}
files <-
  dir_ls(
    path = here("data"),
    type = "file",
    regexp = ".txt$"
  )
```

```{r}
cli_progress_bar(
  total = length(files),
  clear = FALSE
)

for (i in files) {
  i_id <- str_remove(basename(i), "_actigraphy-processed-data.txt")

  i_ga_start <-
    control_data |>
    filter(id == i_id) |>
    pull(ga_start)

  i_data <-
    i |>
    read_acttrust(tz = "America/Sao_Paulo") |>
    shush() |>
    mutate(
      ga =
        ga_point(
          i_ga_start,
          timestamp
        ) |>
        as.numeric() |>
        as.duration() |>
        add(as_hms(timestamp) |> as.numeric() |> as.duration()),
      ga_week =
        ga |>
        as.numeric() |>
        divide_by(as.numeric(ddays(7))) |>
        as.integer()
    )

  i_sri_data |> write_rds(here("data", paste0(i_id, "_sri-data.rds")))

  cli_progress_update()
}

cli_progress_done()
```

## Compute SRI

Critérios de exclusão:

- Semanas com menos de 5 dias
- Semanas com menos de 75% de disponibilidade de dados SRI

Passos:

1. Calcular a IG em cada ponto do dado actigráfico.
2. Remover semanas com menos de 5 dias
3. Calcular o SRI para cada semana (summarize)
4. Transformar em NA SRIs com proporção de valores de concordâncias menores que `0.75`.
5. Remover semanas com menos de 75% de disponibilidade de dados SRI.
6. Tirar a média do SRI por semana (colunas: id, ig, sri)
7. Juntar os dados de todas as gestantes.
8. ANOVA do `sri` por grupos de IG

```{r}
files <-
  dir_ls(
    path = here("data"),
    type = "file",
    regexp = ".txt$"
  )
```

```{r}
cli_progress_bar(
  total = length(files),
  clear = FALSE
)

for (i in files) {
  i_id <- str_remove(basename(i), "_actigraphy-processed-data.txt")

  i_ga_start <-
    control_data |>
    filter(id == i_id) |>
    pull(ga_start)

  i_data <-
    i |>
    read_acttrust(tz = "America/Sao_Paulo") |>
    shush() |>
    mutate(
      ga =
        ga_point(
          i_ga_start,
          timestamp
        ) |>
        as.numeric() |>
        as.duration() |>
        add(as_hms(timestamp) |> as.numeric() |> as.duration()),
      ga_week =
        ga |>
        as.numeric() |>
        divide_by(as.numeric(ddays(7))) |>
        as.integer()
    ) |>
    as_tibble() |>
    group_by(ga_week) |>
    mutate(
      ga_week_available_days = (n() * 60) / as.numeric(ddays(1))
    ) |>
    ungroup(ga_week) |>
    filter(ga_week_available_days >= 5) |>
    as_tsibble(index = timestamp)

  if (nrow(i_data) == 0) {
    cli_alert_info(
      paste0(
        "{.strong {col_red(i_id)}} was dropped because it has no gestational ",
        "week with at least 5 days of data."
      )
    )

    next

    cli_progress_update()
  }

  i_sri_data <- tibble()

  for (j in unique(i_data$ga_week)) {
    j_data <- i_data |> filter(ga_week == j)
    j_data_first_index <- which(i_data$ga_week == j)[1]

    if (j_data_first_index > 1) {
      j_data <-
        i_data |>
        slice(seq(j_data_first_index - 1440, j_data_first_index - 1)) |>
        as_tibble() |>
        bind_rows(j_data |> as_tibble()) |>
        as_tsibble(index = timestamp)
    }

    i_sri_data <-
      i_sri_data |>
      bind_rows(
        j_data |>
          sri(
            sleeping_states = c(1, 2),
            awake_states = 0,
            min_data = 0.75
          ) |>
          mutate(ga_week = j) |>
          as_tibble()
      )
  }

  i_sri_data <-
    i_sri_data |>
    drop_na(sri) |>
    group_by(ga_week) |>
    mutate(dummy = n() >= (1440 * 0.75)) |> # 1080
    ungroup() |>
    filter(dummy)

  if (nrow(i_sri_data) == 0) {
    cli_alert_info(
      paste0(
        "{.strong {col_red(i_id)}} was dropped because it has no gestational ",
        "week with at least 75% of SRI data."
      )
    )

    next

    cli_progress_update()
  }

  i_sri_data <-
    i_sri_data |>
    group_by(ga_week) |>
    summarize(sri = mean(sri, na.rm = TRUE)) |>
    mutate(id = i_id) |>
    select(id, ga_week, sri)

  i_sri_data |> write_rds(here("data", paste0(i_id, "_sri-data.rds")))

  cli_progress_update()
}

cli_progress_done()
```

```{r}
read_rds(here("data", paste0(i_id, "_sri-data.rds")))
```

## Group SRI Data

```{r}
files <-
  dir_ls(
    path = here("data"),
    type = "file",
    regexp = ".rds$"
  )
```

```{r}
sri_data <- tibble()

for (i in files) {
  sri_data <-
    i |>
    read_rds(i) |>
    bind_rows(sri_data)
}

sri_data <-
  sri_data |>
  mutate(ga_week = as.factor(ga_week)) |>
  arrange(ga_week)
```

```{r}
sri_data
```

## Compute State Proportions

```{r}
files <-
  dir_ls(
    path = here("data"),
    type = "file",
    regexp = ".txt$"
  )
```

```{r}
cli_progress_bar(
  total = length(files),
  clear = FALSE
)

for (i in files) {
  i_id <- str_remove(basename(i), "_actigraphy-processed-data.txt")

  i_data <-
    i |>
    read_acttrust(tz = "America/Sao_Paulo") |>
    shush() |>
    state_prop() |>
    mutate(id = i_id)

  i_data |> write_rds(here("data", paste0(i_id, "_state-prop-data.rds")))

  cli_progress_update()
}

cli_progress_done()
```

## Explore Model Data

### SRI by Gestational Age Week

```{r}
sri_data |>
  ggplot(aes(x = sri, fill = ga_week)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    x = "Sleep Regularity Index (SRI)",
    y = "Gestational Age Week"
  )
```

```{r}
sri_data |>
  ggplot(aes(x = sri, y = ga_week, color = ga_week)) +
  geom_point() +
  geom_jitter(width = 0.1, height = 0) +
  labs(
    x = "Sleep Regularity Index (SRI)",
    y = "Gestational Age Week"
  )
```

## SRI Over Time

(all sample combined)

### State Proportions

```{r}
i_data |>
  mutate(per = prop * 100) |>
  ggplot(ggplot2::aes(x = time, y = per)) +
  geom_smooth(color = "#FC2913") +
  labs(
    x = "Time of day (Hour)",
    y = "Percentage of time asleep (%)",
  ) +
  scale_x_time(
    breaks = breaks_width("6 hours"),
    labels = label_time("%-H") # Use "%#H" for Windows
  ) +
  scale_y_continuous(limits = c(NA, 100))
```

## Prepare Model Data for ANOVA

Critério de exclusão:

- Grupos de semanas gestacioanis com menos de 5 gestantes

```
Ho: 36 >= 37 >= 38 >= 39 >= ...
Ha: 36 < 37 < 38 < 39 < ...
```
- Fazer ANCOVA
- Ver outros modelos

```{r}
model_data <-
  sri_data |>
  group_by(ga_week) |>
  mutate(dummy = n() >= 5) |>
  ungroup(ga_week) |>
  filter(dummy) |>
  arrange(ga_week) |>
  select(-dummy)
```

```{r}
model_data |> summarize(n = n(), .by = ga_week)
```

## *A Prioi* Power Analysis

```{r}
pwr_analysis <- pwr.anova.test(
  k = model_data |> pull(ga_week) |> unique() |> length(),
  f = 0.25, # Cohen's f for medium effect sizes
  sig.level = 0.05,
  power = 0.8
)
```

```{r}
pwr_analysis
```

```{r}
pwr_analysis |>
  plot.power.htest(
    xlab = "Sample size per group",
    ylab = "Power (1 - Beta)",
    main = NULL
  )
```

## Model Data

```{r}
observed_statistic <-
  model_data |>
  specify(sri ~ ga_week) |>
  hypothesize(null = "independence") |>
  calculate(stat = "F")

observed_statistic
```

```{r}
null_dist <-
  model_data |>
  specify(sri ~ ga_week) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "F")

null_dist
```

```{r}
null_dist |>
  visualize(method = "both") +
  shade_p_value(
    obs_stat = observed_statistic,
    direction = "greater"
  ) +
  labs(
    title = NULL,
    x = "F-statistic",
    y = "Frequency"
  )
```

```{r}
null_dist |>
  get_p_value(
    obs_stat = observed_statistic,
    direction = "right"
  )
```

```{r}
aov(sri ~ ga_week, data = model_data) |>
  TukeyHSD(conf.level = 0.95) |>
  tidy()
```

## Effect Size

```{r}
effect_size <-
  aov(sri ~ ga_week, data = model_data) |>
  cohens_f()

effect_size
```

```{r}
effect_size[[2]] |>
  f_to_eta2() |>
  interpret_eta_squared(rules = "cohen1992")
```

## *A Posteriori* Power Analysis

```{r}
groups_n <-
  model_data |>
  summarize(n = n(), .by = ga_week)

groups_n
```

```{r}
pwr.anova.test(
  k = nrow(groups_n),
  n = min(groups_n$n),
  f = effect_size[[2]],
  sig.level = 0.05
)
```

## Other

```{r}
file <-
  here::here("data") |>
  list.files(full.names = TRUE) |>
  sample(1)
```

```{r}
# Adicionar no `actverse`

file |>
  readr::read_lines() |>
  iconv(to = "ASCII//TRANSLIT") |>
  readr::write_lines(file)
```

```{r}
data <-
  file |>
  read_acttrust(tz = "America/Sao_Paulo")
```

```{r}
#| eval: false

data |> actogram(days = Inf)
```

```{r}
sri_data <- data |> sri()
```

```{r}
sri_data |>
    ggplot(ggplot2::aes(x = time, y = sri)) +
    # geom_line() +
    geom_smooth(color = "#FC2913") +
    labs(
      x = "Time of day (Hour)",
      y = "Sleep Regularity Index (SRI)"
    ) +
    scale_x_time(
      breaks = breaks_width("6 hours"),
      labels = label_time("%-H") # Use "%#H" for Windows
    ) +
    scale_y_continuous()
```

## Data Modeling

## Hypothesis Testing
