# Welcome {.unnumbered}

```{r}
#| label: setup
#| include: false

library(badger)
library(here)

source(here("R", "_setup.R"))
```

This book represents an collection of data reports related to the study: *Does Sleep Regularity Worsen Over the Third Trimester of Pregnancy?*.

All analyses presented in this document are fully reproducible and were conducted using the [Quarto](https://quarto.org/) publishing system and the [R](https://www.r-project.org/) programming language. Click [here](https://github.com/aliciarvilefortsales/sleep-in-pregnancy) or on the GitHub icon in the menu to explore the code. The research compendium is also accessible by clicking [here](https://doi.org/10.17605/OSF.IO/S4TBZ) via [The Open Science Framework](https://osf.io/).

## License {.unnumbered}

```{r}
#| echo: false
#| output: asis

cat(
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  ),
  badge_license(
    "CC BY-NC-SA 4.0",
    "lightgrey",
    url = "https://creativecommons.org/licenses/by-nc-sa/4.0/"
  )
)
```

The code in this repository is licensed under the [GNU General Public License Version 3](https://www.gnu.org/licenses/gpl-3.0), while the documents are available under the [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-nc-sa/4.0/).

The research data is subject to a private license and is not publicly available due to privacy and ethical considerations.

```text
Copyright (C) 2025 Alícia Rafaelly Vilefort Sales

The code in this repository is free software: you can redistribute it and/or
modify it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or (at your
option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the GNU General Public License for more details.

You should have received a copy of the GNU General Public License along with
this program. If not, see <https://www.gnu.org/licenses/>.
```
