# sleep-in-pregnancy

<!-- quarto render -->

```{r}
#| label: setup
#| include: false

library(badger)
library(here)
library(htmltools)
library(knitr)

source(here("R", "_setup.R"))
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badge_repostatus("active"),
  badge_custom(
    x = "OSF DOI",
    y = "10.17605/OSF.IO/EAPZU",
    color = "1284C5",
    url = "https://doi.org/10.17605/OSF.IO/S4TBZ"
  ),
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  ),
  badge_license(
    "CC BY-NC-SA 4.0",
    "lightgrey",
    url = "https://creativecommons.org/licenses/by-nc-sa/4.0/"
  )
)
```
<!-- badges: end -->

## Overview

This repository contains materials for the study: *Does Sleep Regularity Worsen Over the Third Trimester of Pregnancy?*. The full research compendium, including data and supplementary materials, is available on the Open Science Framework ([OSF](https://osf.io/)) at: <https://doi.org/10.17605/OSF.IO/S4TBZ>

## Keys

To access the data and run the [Quarto](https://quarto.org/) notebooks in the `qmd` folder, you must first obtain authorization to access the project's [OSF](https://osf.io) repositories and [Google Sheets](https://workspace.google.com/products/sheets/) files.

Once you have the necessary permissions, run the following command to authorize your access to the Google Sheets API:

```r
library(gargle)
library(googlesheets4)

options(gargle_oauth_cache = ".secrets")

gs4_auth()
gargle_oauth_cache()
```

Next, create a file named [`.Renviron`](https://bookdown.org/csgillespie/efficientR/set-up.html#:~:text=2.4.6%20The%20.Renviron%20file) in the root directory of the project and add the following environment variables:

- `OSF_PAT`: Your [OSF](https://osf.io/) Personal Access Token ([PAT](https://en.wikipedia.org/wiki/Personal_access_token)). If you don't have one, go to the settings section of your OSF account and create a new token.
- `PREGNANCY_PASSWORD`: The password for the project's [RSA](https://en.wikipedia.org/wiki/RSA_cryptosystem) private key (32 bytes).
- `PREGNANCY_SALT`: The salt used for anonymizing participant IDs (32 bytes).

Example (do not use these values):

```ini
OSF_PAT=bWHtQBmdeMvZXDv2R4twdNLjmakjLUZr4t72ouAbNjwycGtDzfm3gjz4ChYXwbBaBVJxJR
PREGNANCY_PASSWORD=MmXN_od_pe*RdHgfKTaKiXdV7KD2qPzW
PREGNANCY_SALT=iLG*Vu7!2c-KxY_LeQW4D6Kc2aFivvy2
```

Additionally, you will need the following keys in the project's [`_ssh`](_ssh) folder:

- `id_rsa`: The project's private RSA key ([RSA](https://en.wikipedia.org/wiki/RSA_cryptosystem) 4096 bits (OpenSSL)).
- `id_rsa.pub`: The project's public RSA key.

These project's keys are provided to authorized personnel only. If you need access, please contact the author.

## License

```{r}
#| echo: false
#| output: asis

cat(
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  ),
  badge_license(
    "CC BY-NC-SA 4.0",
    "lightgrey",
    url = "https://creativecommons.org/licenses/by-nc-sa/4.0/"
  )
)
```

The code in this repository is licensed under the [GNU General Public License Version 3](https://www.gnu.org/licenses/gpl-3.0), while the documents are available under the [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-nc-sa/4.0/).

The research data is subject to a private license and is not publicly available due to privacy and ethical considerations.

```text
Copyright (C) 2025 Alícia Rafaelly Vilefort Sales

The code in this repository is free software: you can redistribute it and/or
modify it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or (at your
option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the GNU General Public License for more details.

You should have received a copy of the GNU General Public License along with
this program. If not, see <https://www.gnu.org/licenses/>.
```

## Acknowledgments

```{r}
#| echo: false
#| output: asis

tags$table(
  tags$tr(
    tags$td(
      tags$br(),
      tags$p(
        tags$a(
          tags$img(
            src = "images/capes-logo.svg",
            width = 125
          ),
          href = "https://www.gov.br/capes/"
        ),
        align = "center"
      ),
      tags$br(),
      width = "30%"
    ),
    tags$td(
      tags$p(
        "This study was financed by the Coordenação de Aperfeiçoamento de ",
        "Pessoal de Nível Superior - Brazil (",
        tags$a(
          "CAPES",
          href = "https://www.gov.br/capes/",
          .noWS = "outside"
        ),
        ") - Finance Code 001, Grant Number 88887.940089/2024-00."
      ),
      width = "70%"
    )
  )
) |>
  raw_html()
```
