# sleep-in-pregnancy

<!-- quarto render --profile gfm -->

```{r}
#| label: setup
#| include: false

library(badger)
library(beepr)
library(groomr) # github.com/danielvartan/groomr
library(lubridate)
library(readr)
library(stringr)
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_repostatus("active"),
  badger::badge_custom(
    x = "OSF DOI",
    y = "10.17605/OSF.IO/EAPZU",
    color = "1284C5",
    url = "https://doi.org/10.17605/OSF.IO/S4TBZ"
  ),
  badger::badge_license(
    license = "MIT",
    color = "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_license(
    license = "CC BY 4.0",
    color = "lightgrey",
    url = "https://creativecommons.org/licenses/by/4.0/"
  )
)
```
<!-- badges: end -->

## Overview

This repository contains materials for the study: *Associations Between the Duration and Quality of Sleep in Third-Trimester Pregnant Women With the Duration of Labor*. The full research compendium, including data and supplementary materials, is available on the Open Science Framework (OSF) at: [https://doi.org/10.17605/OSF.IO/S4TBZ](https://doi.org/10.17605/OSF.IO/S4TBZ)

## License

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_license(
    license = "MIT",
    color = "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_license(
    license = "CC BY 4.0",
    color = "lightgrey",
    url = "https://creativecommons.org/licenses/by/4.0/"
  )
)
```

The code accompanying this study is distributed under the [MIT License](https://opensource.org/license/mit). All documents are released under the [Creative Commons Attribution 4.0 International Public License](https://creativecommons.org/licenses/by/4.0/).

The research data is subject to a private license and is not publicly available due to privacy and ethical considerations.
